---
title       : Interactive World Map
subtitle    : Population, GDP, and Human Development Index
author      : blackCatMargarita
job         : Courserian
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
license     : by-nc-sa
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
 
---

<style> 
.title-slide {background-color: #E0F8F7; /* #EDE0CF; ; #CA9F9D*/}

.title-slide hgroup > h1{font-family: 'Oswald', 'Helvetica', sanserif;}

.title-slide hgroup > h1, 
.title-slide hgroup > h2 {color: #0B6138;  /* ; #EF5150*/} </style>

## Motivation

* Initial effort to re-create visualizations presented in the Coursera Course, Configuring the World: A Critical Political Economy Approach.  The course is taught by Professor Richard T. Griffiths from Leiden University.  
* The indices chosen for this initial interactive visualization include:
    - Population
    - Gross Domestic Product (GDP) in Purchasing Power Parity (PPP)
    - GDP in current dollars
    - Human Development Index (HDI)
* The data is taken from the databases of the course and sources are explained in the About section of the app.

---

## Summary Statistics

```{r echo = FALSE}
countryData <- read.csv('countries.csv', stringsAsFactors = F)
popData <- countryData[, c('Country','Region', 'IOS','Population','HDI','GDP_PPP','GDP')]
blank <- popData$Population == 'n/a' | popData$HDI == 'n/a' | popData$GDP_PPP == 'n/a' | popData$GDP == 'n/a'
dataMap <- popData[!blank,]
dataMap$Population <- sapply(dataMap$Population, function(x){
    x2 <- gsub(pattern = ",", replacement = "", x = x)
    as.numeric(x2)})
dataMap$GDP_PPP <- sapply(dataMap$GDP_PPP, function(x){
    x2 <- gsub(pattern = ",", replacement = "", x = x)
    as.numeric(x2)})
dataMap$GDP <- sapply(dataMap$GDP, function(x){
    x2 <- gsub(pattern = ",", replacement = "", x = x)
    as.numeric(x2)})
dataMap$HDI <- as.numeric(dataMap$HDI)
````

The following is a summary of the data used in the shiny app found at [https://margarita.shinyapps.io/worldIndices/](https://margarita.shinyapps.io/worldIndices/).  As you will see, the interactive map will demonstrate some advantages over simple plots.

```{r}
summary(dataMap[, c('Population','HDI','GDP_PPP','GDP')])
```

--- 

## GDP and Population

GDP and Population as seen in scatter plot isn't quite as interesting as seeing the data in an interactive map view, but they have their place. Here, we can guess at the outlier countries.

```{r fig.height=5, fig.width=10}
library(ggplot2)
qplot(Population, GDP, data = dataMap, col = Region)
#plot(Population ~ HDI, data = dataMap, color='Region', type='point')
```

---

## Human Development Index

A boxplot summary of HDI by region when no maps are available. 
```{r fig.height=6, fig.width=10}
boxplot(HDI ~ Region, data=dataMap, ylab="HDI")
```

